Optimal (Multiway) Spatial Joins

### 2024-05-20 msj.tex
yufei

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle. Every logarithm has base 2 by default.

@@

\begin{table*} 
    \begin{tabular}{c|c|c|c|c} 
        $\bm{k}$ & {\bf method} & {\bf runtime} & {\bf space} & {\bf remark} \\
        \hline\hline 
        2 & folklore & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal \\ 
        \hline
        $\ge 3$ & before 2022 & $O(n^k)$ & - & \\ 
        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & \\ 
        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & \\ 
        \hline
        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal
    \end{tabular}

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extend the conventional equi-join in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with an ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $k$-tuples $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying $\bigcap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$. It is clear that there is one-one correspondence between the result of this join and that of k-SJ. Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time. Their algorithm requires $O(n \log n)$ space.

@@

\begin{theorem} \label{thm:main-alg}
    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space.
\end{theorem}

>>>>

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle. All logarithms have base 2 by default.

@@

\begin{table*} 
    \begin{tabular}{c|c|c|c|c} 
        $\bm{k}$ & {\bf method} & {\bf runtime} & {\bf space} & {\bf remark} \\
        \hline\hline 
        2 & folklore & $O(n \log n + \out)$ & $O(n)$ & comparison-based optimal \\ 
        \hline
        $\ge 3$ & before 2022 & $O(n^k)$ & - & \\ 
        $\ge 3$ & \cite{ty22} & $O((n + \out) \cdot \polylog n)$ & $O(n \polylog n)$ & $k = O(1)$ \\ 
        $\ge 3$ & \cite{kcko22} & $O(n \log^{2k} n + \out)$ & $O(n \log n)$ & $k = O(1)$ \\ 
        \hline
        $\ge 3$ & ours & $O(n \log n + \out)$ & $O(n)$ & $k = O(1)$, comparison-based optimal
    \end{tabular}

@@

In \cite{kcko22}, Khamis et al.\ investigated a type of joins that extends the conventional equi-join in two ways. First, each attribute value in a relation is an interval (rather than a real value); second, each equality predicate in equi-join is replaced with a ``non-empty intersection'' predicate on the attributes involved. The $k$-SJ problem can be converted to solving a join defined next under the framework of \cite{kcko22}. For each $i \in [k]$, define $R_i$ as a relation over two attributes $X$ and $Y$. For each tuple $\bm{u} \in R_i$, its values $\bm{u}(X)$ and $\bm{u}(Y)$ on the two attributes are both intervals (effectively defining a rectangle). The objective is to output all $k$-tuples $(\bm{u}_1, \bm{u}_2, ..., \bm{u}_k) \in R_1 \times R_2 \times ... \times R_k$ satisfying $\bigcap_{i=1}^k \bm{u}_i(X) \ne \emptyset$ and $\bigcap_{i=1}^k \bm{u}_i(Y) \ne \emptyset$. It is clear that there is one-one correspondence between the result of this join and that of k-SJ. Khamis et al.\ \cite{kcko22} developed an algorithm that can process the join  in $O(n \log^{2k} n + \out)$ time. Their algorithm requires $O(n \log n)$ space.

@@

\begin{theorem} \label{thm:main-alg}
    For $k \ge 3$, we can settle $k$-SJ with a comparison-based algorithm in $O( c^k (k!)^3 \cdot (n \log n + \out))$ time and $O(k \cdot n)$ space, where $c > 1$ is a positive constant.
\end{theorem}
